DynaProg for Scala

### 2013-01-19 docs/report/benchmarks.tex
Fixed one bug, still incorrect.
Nits.

\subsection{Metrics}
The main metrix of interest is the running time. In an attempt to reduce the variance, we run 20 consecutive test and take the median running time (since median is less sensitive to outlier than the average).

@@

>>>>

\subsection{Metrics}
The main metrics of interest is the running time. In an attempt to reduce the variance, we would like to run multiple consecutive test and take the median running time (since median is less sensitive to outlier than the average). Unfortunately, several factors hampers these ideal conditions. First the variance in the running time of CUDA kernels might be quite significant, in particular for short running time. This is due to the fact that the GPU needs to be 'warmed-up' during few seconds before actual computation can happen

  In case the variance is significant, we wou

talk about input sensivity, backup with results on rnafold and matrix mult

@@

XXX: future work: infer automatically "always non-emptiness"

### 2013-01-19 docs/report/implementation.tex
Nits.

% ------------------------------------------------------------------------------------------------
\subsection{LibRNA}
Since the energy computation for RNA folding involve complex coefficient and computations (that are also standardized through libraries and coefficient files), we might want to provide the user with it. To do that, we based our library on \cite{gapc_thesis} which itself is based on \href{http://www.tbi.univie.ac.at/~ivo/RNA/}{ViennaRNA}.\cite{vienna_rna}. Since the library is provided in C, we rely on JNI to reuse the code without modifying it; this allows Scala to immediately benefit from it, but also makes possible to write a GPU version, provided that the related functions are simple enough to be expressible in CUDA.

{\color{red}
XXX: attempt to strip it to make CUDA version
}

>>>>

% ------------------------------------------------------------------------------------------------
\newpage
\subsection{LibRNA}
Since the energy computation for RNA secondary structure prediction (folding the sequence in two dimensions) involve complex coefficients and computations (seemingly standardized in coefficient files), we might want to provide the user with a simple interface to benefit from it. To do that, we based our library on the work of GAPC\cite{gapc_thesis} which itself is based on \href{http://www.tbi.univie.ac.at/~ivo/RNA/}{ViennaRNA}\cite{vienna_rna}. Since the library is provided in C, we rely on JNI to reuse the code without modifying it; this allows Scala to immediately benefit from it, but also makes possible to write a GPU version, provided that the related functions are simple enough to be expressible in CUDA. Our work in this direction is mainly focused on integration, we do not want to discuss the implementation details here but simply give a overview of what we transformed and adapted to suit our needs.

In the first phase, we adapted the library embedded in GAPC to obtain coefficients. Since GAPC is written in C, we had to write a JNI interface to let the Scala code communicate with the libraries. Once this step has been achieved, we focused on obtaining correct results for RNA folding. This has been achieved by a thorough analysis of the GAPC related code, and took quite a long time due to bugs that were hard to find.

In parallel with this work, we focused on making the code compatible with CUDA. We managed to significantly reduce its size by removing unused functions (actually, all the programs of the Vienna package reuse the library, each introducing its own functions). This also enabled us to have a better understanding of the involved computations, thereby helping us to clearly separate the coefficient file processing and the energies computations. We also provided small optimizations towards parallelization and simplified the sequence management (because it needs to be converted in a particular format for the library to efficiently process it).

Once the library was ready to fit on CUDA devices, we integrated it into DynaProg. Unfortunately, since the library adds a significant volume of code (because some coefficients are embedded in C files), the compilation process length was increased (from approx. 2 to 7 seconds). To reduce this penalty (towards benchmarking), we introduced memoization of the compilation results in our code generator, thereby avoiding duplicated compilations of the program for same length of sequences (generated programs are tailored for particular sequences lengths).
